---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uesdRobust

<!-- badges: start -->
<!-- badges: end -->

The uesdRobust R package accompanies the paper “Type I Error Inflation in Unexpected Event during Survey Designs" by Joris Frese and Sascha Riaz. In this paper, we empirically demonstrate a substantially inflated rate of false-positive findings in UESD research and propose ways to adjust UESD test statistics accordingly.

The goal of the uesdRobust R package is to assist users in analyzing (OLS or RDD) UESDs. The package also allows users to estimate placebo-robust p-values and confidence intervals that adjust UESD results to context-specific Type I error inflation. For these purposes, the package offers five complementary functions: **uesd_robust_ols** for (placebo-)robust UESDs with the OLS estimator, **uesd_robust_rdd** for (placebo-)robust UESDs with the RDD estimator, **uesd_placebo_plot** to visualize the empirical distribution of placebo T or Z statistics and the share of placebo tests with more extreme statistics than the main test, **uesd_coef_plot** to visualize the UESD ITT with its unadjusted and placebo-robust 95% confidence interval, and **uesd_table** to create a LaTeX table of the unadjusted and placebo-robust results of multiple estimations.

## Installation

You can install the **uesdRobust** R package like so:

``` r
devtools::install_github("fresej/uesdRobust")
```

## uesd_robust_ols

The **uesd_robust_ols** function is used to analyze UESDs with the OLS estimator. It has 2 basic purposes: **1)** The function wraps features from stats, sandwich, and lmtest to estimate basic dummy-OLS UESDs with heteroskedasticity-robust or cluster-robust standard errors, requiring only a data frame, an outcome variable, and a numeric running variable (i.e., a date variable turned into a numeric indicator) as inputs. We hope that this wrapper makes UESD estimation more accessible for researchers new to this method. **2)** The function additionally allows for the estimation of placebo tests with cutoffs across the whole range of the running variable. Based on these estimations, it creates an empirical distribution of T-statistics that is used to create a placebo-robust p-value and 95% confidence interval.

The function contains the following parameters:

**df**: A data.frame containing your data (mandatory).

**outcome**: String name of the numeric outcome variable (mandatory).

**date_num**: String name of the numeric running date variable (mandatory).

**controls**: NULL (default) or a character vector of other column-names in `df` to add as linear controls.

**bw**: Numeric half‐width of the window around `cutoff`. Defaults to 15.

**cutoff**: Numeric cutoff on `date_num` defining the treated (≥cutoff) vs control (<cutoff) groups. Defaults to 0.

**cluster**: NULL (default) for heteroskedasticity‐robust SE, or a string naming a column in `df` on which to cluster.

**placebos**: Logical; if TRUE, runs regressions at every feasible cutoff (skipping `excluded` after the main), and computes a placebo‐robust p‐value and CI.

**excluded**: Integer; number of subsequent cutoffs after the main one to skip in the placebo loop (default = 3).



## uesd_robust_rdd

The **uesd_robust_rdd** function is used to analyze UESDs with the RDD estimator. It has 2 basic purposes: **1)** The function wraps features from stats and rdrobust to estimate basic RDD UESDs with heteroskedasticity-robust and bias-corrected standard errors, requiring only a data frame, an outcome variable, and a numeric running variable (i.e., a date variable turned into a numeric indicator) as inputs. We hope that this wrapper makes UESD estimation more accessible for researchers new to this method. **2)** The function additionally allows for the estimation of placebo tests with cutoffs across the whole range of the running variable. Based on these estimations, it creates an empirical distribution of Z-statistics that is used to create a placebo-robust p-value and 95% confidence interval.

The function contains the following parameters:

**df**: A data.frame containing your data (mandatory).

**outcome**: String name of the numeric outcome variable (mandatory).

**running**: String name of the numeric running (forcing) variable (mandatory).

**cutoff**: Numeric discontinuity point on `running`. Defaults to 0.

**bw**: Numeric half-width of the RDD window (h). If NULL (default), uses rdrobust’s optimal bandwidth.

**placebos**: Logical; if TRUE, runs placebo RDDs at every feasible cutoff (skipping `excluded` after the main) and computes a placebo-robust p-value and CI.

**excluded**: Integer; number of subsequent cutoffs after the main one to skip in the placebo loop (default = 3).

**se_type**: Character; which rdrobust row to use for placebo inference. One of
    "conventional" (row 1), "bias-corrected" (row 2), or "robust" (row 3).  Default: "conventional".
    
    
## uesd_placebo_plot

The **uesd_placebo_plot** function is used to visualize the empirical distribution of UESD placebo test statistics and the relative position of the original UESD test statistic. It takes the list objects created by **uesd_robust_ols** or **uesd_robust_rdd** as inputs and wraps features from ggplot2 to visualize them.

The function contains the following parameters:

**res**: A list returned (invisibly) by `uesd_robust_ols(..., placebos = TRUE)`
             or `uesd_robust_rdd(..., placebos = TRUE)`. It contains the results of all placebo tests and the main test.
             
**binwidth**: Numeric. Width of the histogram bins (default = 0.1).



## uesd_coef_plot

The **uesd_coef_plot** function is used to visualize the ITT coefficient comparatively with its unadjusted and its placebo-robust 95% confidence interval. It takes the list objects created by **uesd_robust_ols** or **uesd_robust_rdd** as inputs and wraps features from ggplot2 to visualize them.

The function contains the following parameters:

**res**: A list returned by one of the UESD functions with components: `estimate`: numeric, the point estimate, `ci`: named length‐2 numeric for the conventional 95% CI, and `ci_placebo`: named length‐2 numeric for the placebo-robust 95% CI.
             
             
## uesd_table

The **uesd_coef_plot** function is used to create a LaTeX table showing the ITT estimate, its unadjusted 95% CI and p-value, its placebo-robust 95% CI and p-value, the sample size, the estimator, and the standard error type. It takes the list objects created by **uesd_robust_ols** or **uesd_robust_rdd** as inputs and wraps features from knitr to tabularize them.

The function contains the following parameters:

**res** One of: a single result‐list (with `estimate`,`se`,`ci`,`p_value`,`ci_placebo`,`placebo_p_value`,`n_obs`, and `se_type`),
a list of such result‐lists, or a character vector of object names in your calling environment.

**digits** Integer: number of decimal places for non‐p‐value entries (default: 3).

**digits_p** Integer: number of decimal places for p‐value entries (default: 5).
             


## Example

This is a basic example using simulated data to demonstrate one hypothetical use case for the **uesdRobust** package.


```{r }
library(uesdRobust)


# simulate some survey data with placebo effects
set.seed(123)
n <- 3000
df_sim <- data.frame(
  running = sample(-150:100, size = n, replace = TRUE)
)
main_cut    <- 0
main_eff    <- 1
placebo_cuts <- c(-76, -66, -56, -39, -30, -25, - 20, -17, -14, -11, -10, 
                  -7, -2,  10, 15, 18, 23, 27, 29,  30, 33, 31, 34,45,66,76)
placebo_eff  <- c( 1.5, -2, 1.9, 1.6, -0.3, 1.7, -0.4,  0.5, -2.3, 0.7, -0.4,  0.5, 
                   -0.3, 0.7, -0.4, 0.7, 0.9, 1.8, 3, 2.1, 2.2, 1.7, 1.1, 1.2, 1, -1)
df_sim$outcome <- 0
df_sim$outcome <- df_sim$outcome +
  main_eff * (df_sim$running >= main_cut)
for(i in seq_along(placebo_cuts)) {
  df_sim$outcome <- df_sim$outcome +
    placebo_eff[i] * (df_sim$running >= placebo_cuts[i])
}
df_sim$outcome <- df_sim$outcome +
  rnorm(n, mean = 0, sd = 1)


# RDD‐based UESD with the uesd_robust_rdd function:
rdd_res <- uesd_robust_rdd(
  df       = df_sim,
  outcome  = "outcome",
  running  = "running",
  cutoff   = 0,
  bw       = NULL,               
  placebos = TRUE,
  se_type  = "conventional"
)


# plot of the placebo distribution of z‐values with the uesd_placebo_plot function:
uesd_placebo_plot(rdd_res)



# plot of the ITT coefficient with its unadjusted and placebo-robust 95% confidence interval:
uesd_coef_plot(rdd_res)


```



